---
title: "Lab 3 -  MplusAutomation, what's missing?"
author: "Factor Analysis ED 216B - Instructor: Karen Nylund-Gibson"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document: default
  pdf_document: default
subtitle: '**Adam Garber**'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, message = FALSE, warning = FALSE, tidy = TRUE)
```

<style type="text/css">
body{ font-size: 20px; max-width: 1600px; margin: auto; padding: 1em; }
code.r{ font-size: 18px; }
p { padding-top: 10px; padding-bottom: 10px; }
pre { font-size: 16px; }
</style>

## ____________________________________

## Getting started - following the routine...

1. Create an R-Project
2. Install packages ($\color{red}{\text{ONLY IF NEEDED}}$)
3. Load packages

### R-Project instructions:

1. click "NEW PROJECT" (upper right corner of window)
2. choose option "NEW DIRECTORY"
3. choose location of project (on desktop OR in a designated class folder)

Within R-studio under the files pane (bottom right):

1. click "New Folder" and name folder "data"
2. click "New Folder" and name folder "efa_mplus"
3. click "New Folder" and name folder "figures"

## ____________________________________


## IF package rhdf5 does not load then run:

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE)) 
  install.packages("BiocManager")

BiocManager::install("rhdf5")
```

# ____________________________________

## loading packages...

```{r, eval=TRUE}
library(apaTables)
library(reshape2)
library(MplusAutomation)
library(rhdf5)
library(tidyverse)
library(here)
library(kableExtra)
```

## Lab 3 - MplusAutomation, what's missing?

**Learning goal:** The goal of this lab activity is to locate areas of the MplusAutomation code which will change depending on the particular data & modeling context.

### Instructions:

- fill in the missing code whenever you see (***)
- speak to your neighbors
- ask questions 
- errors are ok & will give us insight into how to fix things
- do this exercize *without refrencing previous labs*
- repetition is key, this is a safe space to learn to see where code might be broken

### Read csv data file into your R environment:

```{r}

lab_data <- read_***(***("***", "***")) # What's missing?

```

## make an apa style table of summary statistics:
https://cran.r-project.org/web/packages/apaTables/vignettes/apaTables.html

## I also like the package "stargazer" for this, see here:
https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf

```{r}
# correlation table with means & SD
apa.cor.table(lab_data[,1:5], filename=here("figures", "Table_cor_lab3_APA.doc"), table.number=1)

```


## Subset data

```{r}

# make a subset, keep all columns except (BYRACE & BYSTLANG)

schl_safe *** lab_data ***     # What's missing?
  ***( )


```

# ____________________________________


### Reverse indicators so scale has consistent meaning for factor interpretation

```{r}
# Reverse code the following columns 1-3, 5-7, 15-19 (by using collumn numbers)

cols = c(***)       # What's missing?

# Reverse coding requires taking the range + 1 ...

schl_safe[,***] <-  *** - schl_safe[ ,***]        # What's missing?

```

# ____________________________________


## Run an efa model

1. with 1 through 4 factors
2. the robust maximum likelihood estimator
3. for **female** observations only (BYSEX = 2)

```{r}
## efa (indicators: school climate, safety, clear rules)

# What's missing?

m_efa_1  <- mplusObject(
  TITLE = "FACTOR ANALYSIS EFA - LAB 2 DEMO", 
  VARIABLE = 
   "! within mplus you can choose a number of sequential columns 
    ! using (var1_name - var_last_name)
     
    usevar = BYS20A-***;
  
    *** = *** == ***;", 
  
  ANALYSIS = 
 "type = *** ***;   
  estimator = ***;  
  parallel=50; ! run the parallel analysis for viewing with the eigenvalue elbow 
  ",
 
  MODEL = "" ,
  
  PLOT = "type = plot3;",
  OUTPUT = "sampstat;",
  
  usevariables = colnames(***), 
  rdata = ***)

m_efa_1_fit <- mplusModeler(***, 
                            dataout=here("***", "***.dat"),
                            modelout=here("***", "***.inp"),
                            check=TRUE, run = TRUE, hashfilename = FALSE)

## END: EXPLORATORY FACTOR ANALYSIS
```

# ____________________________________
# ____________________________________

## Run an efa with reduced set of items (see comments "!" below )

### This time with...

- reduced set of items (see comments "!" below about items to exclude)
- 1 through 5 factors 
- the robust maximum likelihood estimator
- parallel analysis
- again, use female observations in the analysis

```{r}
## efa reduced set - What's missing? 

m_efa_1  <- *** (
  TITLE = "FACTOR ANALYSIS EFA - REDUCED SET - LAB 2 DEMO", 
  VARIABLE = 
    "usevar =  
     ***
     ! remove: BYS20C BYS20D
     ***
     ! remove:BYS20H BYS20I BYS20L
     ***
     ! remove: BYS21B 
     ;",
  
  ANALYSIS = 
    " *** = *** ***   
      *** = ***   
      *** = ***        
    ",
  
  MODEL = "" ,
  
    PLOT = "*** = ***;",
  OUTPUT = "***;",
  
  usevariables = colnames(***), 
  rdata = ***)

m_efa_2_fit <- mplusModeler(***, 
                            dataout=here("***", "***.dat"),
                            modelout=here("***", "***.inp"),
                            check=TRUE, run = TRUE, hashfilename = FALSE)

## END: EXPLORATORY FACTOR ANALYSIS OF - REDUCED SET
```

## Make a tribble table of your factor loadings ...

```{r, eval=TRUE}

loading_table <- tribble(
  ~"Items", ~"Factor 1", ~"Factor 2",  ~"Factor 3",
 #----------|-------------|------------|-----------|,
  "***"  ,  "***"  ,  "***"  ,  "***"  ,   
  "***"  ,  "***"  ,  "***"  ,  "***"  ,    
  "***"  ,  "***"  ,  "***"  ,  "***"  ,    
  "***"  ,  "***"  ,  "***"  ,  "***"  ,    
  "***"  ,  "***"  ,  "***"  ,  "***"  ,    
  "***"  ,  "***"  ,  "***"  ,  "***"  ,
)

loading_table %>% 
  kable() %>% 
  kable_styling(latex_options = c("striped"), 
                full_width = F,
                position = "left")
```

# ____________________________________

## Plot Parallel Analysis & Eigenvalues

### read into R an Mplus output file

```{r}
efa_summary <- readModels(here("efa_mplus", "lab3_efa2_female.out"))
```

### extract relavent data & prepare dataframe for plot

```{r}
x <- list(EFA=efa_summary[["gh5"]][["efa"]][["eigenvalues"]], 
          Parallel=efa_summary[["gh5"]][["efa"]][["parallel_average"]])

plot_data <- as_data_frame(x)
plot_data <- cbind(Factor = paste0(1:nrow(plot_data)), plot_data)

plot_data <- plot_data %>% 
  mutate(Factor = fct_inorder(Factor))
```

### pivot the dataframe to "long" format

```{r}
plot_data_long <- plot_data %>% 
  pivot_longer(EFA:Parallel,               # The columns I'm gathering together
               names_to = "Analysis",      # new column name for existing names
               values_to = "Eigenvalues")  # new column name to store values
```

### plot using ggplot

```{r}
plot_data_long %>% 
ggplot(aes(y=Eigenvalues,
           x=Factor,
           group=Analysis,
           color=Analysis)) +
  geom_point() + 
  geom_line() + 
  theme_minimal()
```

### save figure to the designated folder

```{r}
ggsave(here("figures", "eigenvalue_elbow_rplot.png"), dpi=300, height=5, width=7, units="in")
```

# ____________________________________

## End of lab 3 exercise

# ____________________________________

## References: 

Hallquist, M. N., & Wiley, J. F. (2018). MplusAutomation: An R Package for Facilitating Large-Scale Latent Variable Analyses in Mplus. Structural equation modeling: a multidisciplinary journal, 25(4), 621-638.

