---
title: "Lab 4 - Random splits, Iterators, & Handling large data"
author: "Factor Analysis ED 216B - Instructor: Karen Nylund-Gibson"
date: "1/28/2020"
output:
  html_document: default
  pdf_document: default
subtitle: 'Adam Garber'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, message = FALSE, warning = FALSE, tidy = TRUE)
```

## Lab 4 outline

1. Randomly split data into 2 equal parts (calibration & validation samples)
2. Introduction to MplusAutomation with iterators
3. Dealing with large data

## ____________________________________

## Getting started - following the routine...

1. Create an R-Project
2. Install packages ($\color{red}{\text{ONLY IF NEEDED}}$)
3. Load packages

### R-Project instructions:

1. click "NEW PROJECT" (upper right corner of window)
2. choose option "NEW DIRECTORY"
3. choose location of project (on desktop OR in a designated class folder)

Within R-studio under the files pane (bottom right):

1. click "New Folder" and name folder "data"
2. click "New Folder" and name folder "efa_mplus"
3. click "New Folder" and name folder "figures"

## ____________________________________

## SKIP (return here if you recieve an $\color{red}{\text{error}}$ when loading packages)

1. First check if the packages load
2. If an error is returned for package(s) 
3. then run the following lines of code
4. AFTER ANY INSTALLATION... MUST RE-LOAD PACKAGES! 

###   ** $\color{red}{\text{NOTE:}}$ the following code ONLY needs to be run once per computer!

```{r}
# to install one package at a time...
install.packages("janitor")
install.packages("haven")

# to install multiple pacakages at once...
install.packages(c("janitor",
                   "MplusAutomation",
                   "tidyverse",
                   "here",
                   "haven",
                   "corrplot"))
```

## IF package rhdf5 does not load then run:

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE)) 
  install.packages("BiocManager")

BiocManager::install("rhdf5")
```

# ____________________________________

## Lab 4 - Begin 

# ____________________________________

### loading packages...

```{r, eval=TRUE}
library(janitor)
library(tidyverse)
library(haven)
library(MplusAutomation)
library(rhdf5)
library(here)
library(corrplot)
```

### read in the raw dataset 

```{r}
lab_data <- read_spss(here("data", "lab_efa_cfa_data.sav"))
```

## create a subset of the dataset called "school_trouble"
```{r}
school_trouble <- lab_data %>% 
  select(41:55)
```

## make a new codebook from the "school_trouble" subset
```{r}
sjPlot::view_df(school_trouble)
```

### write a CSV datafile
```{r}
write_csv(school_trouble, here("data", "school_trouble_data.csv"))
```

### read the unlabeled data back into R
```{r}
trouble_data <- read_csv(here("data", "school_trouble_data.csv"))
```

### check items to see if reverse coding is needed 
```{r}
cor_matrix <- cor(trouble_data, use = "pairwise.complete.obs")

corrplot(cor_matrix, method="circle",
         type = "upper")
```

# ____________________________________
#
## Randomly split a sample into 2 equal parts
#
# ____________________________________

### find the size of half of original sample. 
### The "floor()" function helps with rounding
```{r}
smp_size <- floor(0.50 * nrow(trouble_data))
```

### set the seed to make your partition reproducible
```{r}
set.seed(123)
```

### the function "sample()" will pick at random the values of the specified number
```{r}
calibrate_smp <- sample(seq_len(nrow(trouble_data)), size = smp_size)
```

### create two samples called "calibrate" & "validate"
```{r}
calibrate <- trouble_data[calibrate_smp, ]
validate <- trouble_data[-calibrate_smp, ]
```

### Let's run an EFA with the "calibrate" sample
```{r}
m_efa_1  <- mplusObject(
  TITLE = "School Trouble EFA - LAB 4 DEMO", 
  VARIABLE = 
    "usevar = BYS22A-BYS24G;", 
  
  ANALYSIS = 
    "type = efa 1 5;   
     estimator = mlr;  
     parallel=50; ! run parallel analysis",
  
  MODEL = "" ,
  
  PLOT = "type = plot3;",
  OUTPUT = "sampstat;",
  
  usevariables = colnames(calibrate), 
  rdata = calibrate)

m_efa_1_fit <- mplusModeler(m_efa_1, 
                            dataout=here("efa_mplus", "lab4_efa1_trouble.dat"),
                            modelout=here("efa_mplus", "lab4_efa1_trouble.inp"),
                            check=TRUE, run = TRUE, hashfilename = FALSE)
```

### Plot Parallel Analysis & Eigenvalues

### read into R an Mplus output file
```{r}
efa_summary <- readModels(here("efa_mplus", "lab4_efa1_trouble.out"))
``` 

### extract relavent data & prepare dataframe for plot
```{r}
x <- list(EFA=efa_summary[["gh5"]][["efa"]][["eigenvalues"]], 
          Parallel=efa_summary[["gh5"]][["efa"]][["parallel_average"]])

plot_data <- as_data_frame(x)
plot_data <- cbind(Factor = paste0(1:nrow(plot_data)), plot_data)

plot_data <- plot_data %>% 
  mutate(Factor = fct_inorder(Factor))
```

### pivot the dataframe to "long" format
```{r}
plot_data_long <- plot_data %>% 
  pivot_longer(EFA:Parallel,               # The columns I'm gathering together
               names_to = "Analysis",      # new column name for existing names
               values_to = "Eigenvalues")  # new column name to store values
``` 

### plot using ggplot
```{r}
plot_data_long %>% 
  ggplot(aes(y=Eigenvalues,
             x=Factor,
             group=Analysis,
             color=Analysis)) +
  geom_point() + 
  geom_line() + 
  theme_minimal()
``` 

### save figure to the designated folder
```{r}
ggsave(here("figures", "eigenvalue_elbow_rplot.png"), dpi=300, height=5, width=7, units="in")
```

# ____________________________________
#
## Introduction to MplusAutomation with iterators 
#
# ____________________________________

### Alternate way to run an EFA with the "calibrate" sample
```{r}
m_efa  <- lapply(1:5, function(k) {
  m_efa2  <- mplusObject(
    TITLE = "School Trouble EFA - LAB 4 DEMO", 
    VARIABLE = 
      "usevar = BYS22A-BYS24G;", 
    
    ANALYSIS = 
      paste("type=efa", k, k), 
    
    MODEL = "" ,
    
    PLOT = "type = plot3;",
    OUTPUT = "sampstat;",
    
    usevariables = colnames(calibrate), 
    rdata = calibrate)
  
  m_efa_2_fit <- mplusModeler(m_efa2, 
                              dataout=sprintf(here("efa_mplus2", "efa_trouble.dat"), k),
                              modelout=sprintf(here("efa_mplus2", "efa_%d_trouble.inp"), k),
                              check=TRUE, run = TRUE, hashfilename = FALSE)
})
```


# ____________________________________
#
## Cleaning & subsetting large datasets 
#
# ____________________________________

### reading SPSS files is a lot slower than reading CSV formatted files
```{r}
hsls_raw <- read_spss(here("data", "hsls_16_student_sub_v1.sav"))
```

### make all column names "lower_snake_case" style
```{r}
hsls_tidy <- hsls_raw %>% 
  clean_names()
```

### select usig the starts_with() function
```{r}
hsls_x1 <- hsls_tidy %>% 
  select(starts_with("x1")) # columns with first 2 characters "x1"
```

### select using the end_with() function
```{r}
hsls_not_sex <- hsls_tidy %>% 
  select(!ends_with("sex")) # columns that do NOT end with "sex"
```

### select using the end_with() function
```{r}
hsls_science <- hsls_tidy %>% 
  select(contains("sci")) # columns that contain characters "sci"

hsls_math <- hsls_tidy %>% 
  select(contains(c("mth" , "math"))) # columns that contain "mth" or "math"
```

### combine different select() arguements
```{r}
hsls_math_sci <- hsls_tidy %>% 
  select(contains(c("mth" , "math", "sci"))) %>%
  select(!starts_with("x1")) %>% 
  select(!ends_with("sex"))
```

### remove characters from the variable names that are greater than 8 characters 
```{r}
names(hsls_math_sci) = str_sub(names(hsls_math_sci), 1, 8)
```

### check if culumn names are unique
```{r}
test.unique <- function(df) { ## function to identify unique columns
  
  length1 <- length(colnames(df))
  length2 <- length(unique(colnames(df))) 
  if (length1 - length2 > 0 ) {
    
  print(paste("There are", length1 - length2, " duplicates", sep=" ")) 
    }
} 

test.unique(hsls_math_sci)
```

### locate duplicates (this will find the column of the first duplicate)
```{r}
anyDuplicated(colnames(hsls_math_sci))

names(hsls_math_sci)
```


### Other functions to consider from the "stringr" package (part of tidyverse): 

- str_remove()
- str_replace() # replace one string pattern with another
- str_match()
- str_pad()     # to remove spaces 
- str_count()
- str_detect()
- str_dup()
- str_extract_all()

# ____________________________________

## End of lab 4 exercise

# ____________________________________

## References: 

Hallquist, M. N., & Wiley, J. F. (2018). MplusAutomation: An R Package for Facilitating Large-Scale Latent Variable Analyses in Mplus. Structural equation modeling: a multidisciplinary journal, 25(4), 621-638.

