---
title: "Lab 5 -  Path Diagrams & EFA rotation"
author: "Factor Analysis ED 216B - Instructor: Karen Nylund-Gibson"
date: "2/3/2020"
output:
  html_document: default
  pdf_document: default
subtitle: 'Adam Garber'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, message = FALSE, warning = FALSE, tidy = TRUE)
```

<style type="text/css">
body{ font-size: 20px; max-width: 1600px; margin: auto; padding: 1em; }
code.r{ font-size: 18px; }
p { padding-top: 10px; padding-bottom: 10px; }
pre { font-size: 16px; }
</style>

## Lab 5 outline

1. EFA rotations, override the default
2. Specify a confirmatory factor analysis (CFA)
3. Generate a path diagram of your CFA model

## ____________________________________

## Getting started - following the routine...

1. Create an R-Project
2. Install packages 
3. Load packages

### R-Project instructions:

1. click "NEW PROJECT" (upper right corner of window)
2. choose option "NEW DIRECTORY"
3. choose location of project (on desktop OR in a designated class folder)

Within R-studio under the files pane (bottom right):

1. click "New Folder" and name folder "data"
2. click "New Folder" and name folder "efa_mplus"
3. click "New Folder" and name folder "cfa_mplus"

```{r}

install.packages("semPlot")
install.packages("stargazer")

# Install previously used pacakages (only if your on a new computer!)
install.packages(c("semPlot",
                   "stargazer",
                   "MplusAutomation",
                   "tidyverse",
                   "here"))
```

# ____________________________________

## Lab 5 - Begin 

# ____________________________________

loading packages...
```{r, eval=TRUE}
library(tidyverse)
library(MplusAutomation)
library(rhdf5)
library(here)
library(semPlot)
library(stargazer)
```

read in data
```{r, eval=TRUE}

trouble_data <- read_csv(here("Lab5_FA", "data", "school_trouble_data.csv"))

```

take a look at the EFA data (same indicators used for lab 4)
```{r}

stargazer(as.data.frame(trouble_data), type="text", digits=1)

```

```{r, echo=FALSE} 
# results='asis', eval=TRUE, 

stargazer(as.data.frame(trouble_data), type="latex",

header = FALSE, digits=1)

```

replicate the SAME random split as last week (do not change the seed)

- get the size of half of original sample
- `floor()` helps with rounding

```{r}
smp_size <- floor(0.50 * nrow(trouble_data))
```

set the seed to make partition reproducible
```{r}
set.seed(123)
```

`sample()` randomly selects rows of size = "smp_size"
```{r}
calibrate_smp <- sample(seq_len(nrow(trouble_data)), size = smp_size)
```

create two samples called "calibrate" & "validate"
```{r}
calibrate <- trouble_data[calibrate_smp, ]
validate <- trouble_data[-calibrate_smp, ]
```

reorder variables & remove column number 15 
```{r}

# removed BYS24G (due to low loading < .2)
calibrate <-  calibrate %>% 
  select(1,3:14,2) 

# do the same for the validate sample
validate <-  validate %>% 
  select(1,3:14,2) 
```

save the calibrate & validate samples using "write_csv"
```{r}
# save calibrate sample
write_csv(calibrate, here("data", "calibrate_trouble_sample.csv"))

# save validate sample
write_csv(validate, here("data", "validate_trouble_sample.csv"))
```

### Default rotation: Geomin Oblique

#### run an EFA with the "calibrate" sample
```{r}
efa_geomin  <- mplusObject(
  TITLE = "Geomin Oblique EFA - LAB 5 DEMO", 
  VARIABLE = 
    " ! removed BYS24G (due to low loading < .2)
    usevar = BYS22A-BYS22B;", 
  
  ANALYSIS = 
    "type = efa 1 3;   
     estimator = mlr;
     rotation = geomin; ! this is the default (added to be explicit)
     parallel=50; ! run parallel analysis",
  
  MODEL = "" ,
  
  PLOT = "type = plot3;",
  OUTPUT = "sampstat;",
  
  usevariables = colnames(calibrate), 
  rdata = calibrate)

efa_geomin_fit <- mplusModeler(efa_geomin, 
                            dataout=here("efa_mplus", "lab5_efa_geomin_oblique.dat"),
                            modelout=here("efa_mplus", "lab5_efa_geomin_oblique.inp"),
                            check=TRUE, run = TRUE, hashfilename = FALSE)
```

### Rotation: Varimax Orthogonal

```{r}
efa_varimax  <- mplusObject(
  TITLE = "Varimax Orthogonal EFA - LAB 5 DEMO", 
  VARIABLE = 
    " ! removed BYS24G (due to low loading < .2)
    usevar = BYS22A-BYS22B;", 
  
  ANALYSIS = 
    "type = efa 1 3;   
     estimator = mlr;
     rotation = varimax; ! orthogonal (no factor correlations)
     parallel=50; ! run parallel analysis",
  
  MODEL = "" ,
  
  PLOT = "type = plot3;",
  OUTPUT = "sampstat;",
  
  usevariables = colnames(calibrate), 
  rdata = calibrate)

efa_varimax_fit <- mplusModeler(efa_varimax, 
                            dataout=here("efa_mplus", "lab5_efa_varimax_orthogonal.dat"),
                            modelout=here("efa_mplus", "lab5_efa_varimax_orthogonal.inp"),
                            check=TRUE, run = TRUE, hashfilename = FALSE)
```

### Estimate a Confirmatory Factor Analysis (CFA) model

```{r}
cfa_validate  <- mplusObject(
  TITLE = "Geomin Oblique EFA - LAB 5 DEMO", 
  VARIABLE = 
    "usevar = BYS22A-BYS22B;", 
  
  ANALYSIS = 
    "estimator = mlr;",
  
  MODEL = 
    "FACTOR_1 by BYS22A BYS22C BYS22D BYS22E BYS22F BYS22G BYS22H;
  
     FACTOR_2 BY BYS24A BYS24B BYS24C BYS24D BYS24E BYS24F BYS22B;" ,
  
  PLOT = "type = plot3;",
  OUTPUT = "sampstat standardized residual modindices (3.84);",
  
  usevariables = colnames(validate), 
  rdata = validate)

cfa_val_fit <- mplusModeler(cfa_validate, 
                            dataout=here("cfa_mplus", "lab5_cfa_validate.dat"),
                            modelout=here("cfa_mplus", "lab5_cfa_validate.inp"),
                            check=TRUE, run = TRUE, hashfilename = FALSE)
```

#### create a path diagram of the CFA model

```{r}

# Read in the model to R within the "cfa_mplus" folder
cfa_output <- readModels(here("cfa_mplus"))

# Plot model:
semPaths(cfa_output,
         intercepts=FALSE,
         fixedStyle = c(1))
```

play with some formatting
```{r}

# Plot model:
semPaths(cfa_output, "std",
         intercepts=FALSE,
         fixedStyle = c(1),
         fade = FALSE,
         color= list(lat = c("light blue"," light green")),
        )

```

## ____________________________________

### End of Lab 5

## ____________________________________









